---
title: "propensity models"
author: "Andre Hiwatig"
date: "2025-05-26"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load Packages

```{r}
library(MatchIt)
library(tableone)
library(ggplot2)
library(readxl)
library(cobalt)
library(gridExtra)
library(gridExtra)

```

## Explore Data

```{r}
# df <- read.csv("new_updated.csv") # from github
# colnames(df) <- gsub("[[:punct:]]", "", colnames(df))
# 
# sum(is.na(df["GenderMale1Female2"]))
# 
# df['TarsusresectedYesNo'] <- sapply(df[['TarsusresectedYesNo']], function(x) {ifelse(x == "Y", 1, 0)})
# 
# sum(is.na(df$TissueResectedmm))
# 
# df$X %in% c(156, 157, 158)
# 
# df[df$X1eyetarsus1eyenormal == "yes", ]
# 
# # View(df)
# 
# df$TarsusresectedYesNo <- factor(df$TarsusresectedYesNo, levels = c(0, 1), labels = c("No", "Yes"))
```

### Original Dataset

```{r}
# excel_sheets("MMCR Data Stats Class.xlsx")
# patient_all <- read_excel("MMCR Data Stats Class.xlsx", sheet = "1999 to 2018")
# primary_ou <- read_excel("MMCR Data Stats Class.xlsx", sheet = "Primary OU")
# primary_excluding <- read_excel("MMCR Data Stats Class.xlsx", sheet = "Primary excluding")
# 
# # Check Rows
# nrow(primary_ou)
# nrow(primary_excluding)
# 
# colnames(patient_all)
# colnames(primary_ou)
# colnames(primary_excluding)
# 
# # primary_excluding_no_last <- primary_excluding[, 1:(ncol(primary_excluding) - 1)]
# # primary_sorted <- primary_excluding_no_last[order(primary_excluding_no_last[["#"]]), ]
# # identical(primary_ou, primary_excluding_no_last)
# 
# # Prepare for Merge Create Subset 
# patient_all_subset <- patient_all[, c('#', 'Patient_name', 'DOB', 'Previous Surgery? (yes =1, No = 0)',
#                                       'Post MMCR Surgery- ptosis?', 'Levator function', 'Sx Date')]
# 
# # Drop NA rows from dataset
# primary_ou <- primary_ou[-(513:517), ]

```


### Duplicated Patient Name Rows

```{r}
# # Step 1: Find names that appear more than once
# dupes <- df$Patientname[duplicated(df$Patientname)]
# 
# # Step 2: Filter all rows where Patientname is one of the duplicates
# df[df$Patientname %in% dupes, ]
# 
# 
# ### Remove Kenneth Low??
```


## Cj Data Cleaning

```{r}
# Read Dataset
df <- read.csv("new_updated.csv")

# Make Column Names Better
colnames(df) <- gsub("[[:punct:]]", "", colnames(df))

# Change Tarsus Ressected Column to Boolean Value
# df['TarsusresectedYesNo'] <- sapply(df[['TarsusresectedYesNo']], function(x) {ifelse(x == "Y", 1, 0)})

# CJ CHANGE: Make into a factor
df$TarsusresectedYesNo <- factor(df$TarsusresectedYesNo, levels = c("N", "Y"), labels = c("No", "Yes"))


# CHANGED: Changed Gender to Categorical: Male and Female
df['Gender'] <- sapply(df[['GenderMale1Female2']], function(x) {ifelse(x == 1, "Male", "Female")})

# Gender Imputation: Mode
get_mode <- function(x) {                       ## Helper Function
  ux <- na.omit(unique(x))
  ux[which.max(tabulate(match(x, ux)))]
}
                                                
mode_value <- get_mode(df[['Gender']])            ## Impute Gender
df$Gender[is.na(df$Gender)] <- mode_value


# Tissue Ressected Imputation: Mean
df$TissueResectedmm[is.na(df$TissueResectedmm)] <- mean(df$TissueResectedmm, na.rm = TRUE)

# Exclude Patients with Surgeries in both eyes
df <- df[!(df$X %in% c(156, 157, 158)), ]

df$GenderMale1Female2 <- as.factor(df$GenderMale1Female2)
df$ODOS <- as.factor(df$ODOS)



head(df)
```

### Explore Age to make cutoffs

```{r}
summary(df$Age)
hist(df$Age, breaks = 20, main = "Age Distribution", xlab = "Age")
boxplot(df$Age, main = "Boxplot of Age", horizontal = TRUE)
```

```{r}
df$age_group <- cut(df$Age, breaks = 3, labels = c("Young", "Middle-aged", "Older"))
table(df$age_group)

# df$age_group <- cut(df$Age,
#                     breaks = c(0, 50, 70, Inf),
#                     labels = c("Young", "Middle-aged", "Older"))

table(df$age_group)

```


## Cj Propensity Models

### Propensity Models (Ratio 1:1) But Beter lol

```{r}
m_nearest <- matchit(TarsusresectedYesNo ~ PreopMRD1 + Age + ODOS,
                     data = df, method = "nearest", ratio = 1)

m_optimal <- matchit(TarsusresectedYesNo ~ PreopMRD1 + Age + ODOS,
                     data = df, method = "optimal", ratio = 1)

m_full <- matchit(TarsusresectedYesNo ~ PreopMRD1 + Age + ODOS,
                  data = df, method = "full")

m_quick <- matchit(TarsusresectedYesNo ~ PreopMRD1 + Age + ODOS,
                   data = df, method = "quick")

m_cem <- matchit(TarsusresectedYesNo ~ PreopMRD1 + Age + ODOS,
                   data = df, method = "cem")

```

### Plot Covariate Models

```{r}
# Plot balance for each method
love.plot(m_nearest, stats = "mean.diffs", abs = TRUE, 
          main = "Covariate Balance: Nearest Matching", threshold = 0.1)

love.plot(m_optimal, stats = "mean.diffs", abs = TRUE, 
          main = "Covariate Balance: Optimal Matching", threshold = 0.1)

love.plot(m_full, stats = "mean.diffs", abs = TRUE, 
          main = "Covariate Balance: Full Matching", threshold = 0.1)

love.plot(m_quick, stats = "mean.diffs", abs = TRUE, 
          main = "Covariate Balance: Quick Matching", threshold = 0.1)

love.plot(m_cem, stats = "mean.diffs", abs = TRUE, 
          main = "Covariate Balance: CEM Matching", threshold = 0.1)

```

```{r}
p1 <- love.plot(m_nearest, stats = "mean.diffs", abs = TRUE,
                threshold = 0.1, stars = "std",
                title = "Nearest")

p2 <- love.plot(m_optimal, stats = "mean.diffs", abs = TRUE,
                threshold = 0.1, stars = "std",
                title = "Optimal")

p3 <- love.plot(m_full, stats = "mean.diffs", abs = TRUE,
                threshold = 0.1, stars = "std",
                title = "Full")

p4 <- love.plot(m_quick, stats = "mean.diffs", abs = TRUE,
                threshold = 0.1, stars = "std",
                title = "Quick")

p5 <- love.plot(m_cem, stats = "mean.diffs", abs = TRUE,
                threshold = 0.1, stars = "std",
                title = "CEM")

gridExtra::grid.arrange(p1, p2, p3, p4, p5, nrow = 3)
```



### Propensity Models (Ratio 1:1) CJ OLD  (SHOWS P-value)

```{r}
# 
# df$TarsusresectedYesNo <- factor(df$TarsusresectedYesNo)
# 
# is.factor(df$TarsusresectedYesNo)
# levels(df$TarsusresectedYesNo)
```


```{r}
## Method = "nearest"

# only included vars. with complete data 
# match_model <- matchit(TarsusresectedYesNo ~ PreopMRD1 + Age + Gender + ODOS,
#                        data = df,
#                        method = "nearest",
#                        ratio = 1)

match_model <- matchit(TarsusresectedYesNo ~ PreopMRD1 + ODOS + Age,
                       data = df,
                       method = "nearest",
                       ratio = 1)

summary(match_model)

# matched dataset
matched_data <- match.data(match_model)
# matched_data[matched_data['subclass'] == 1, ] # example

treated <- matched_data[matched_data$TarsusresectedYesNo == "Yes", ]
control <- matched_data[matched_data$TarsusresectedYesNo == "No", ]

# Nonparametric Test: Wilcox test
wilcox.test(MostrecentPostOpMRD1 ~ TarsusresectedYesNo, data = matched_data)
wilcox.test(treated$MostrecentPostOpMRD1, control$MostrecentPostOpMRD1, paired = TRUE)


```

```{r}
## Method = "optimal"

# df$ODOS <- factor(df$ODOS)

# only included vars. with complete data 
match_model <- matchit(TarsusresectedYesNo ~ PreopMRD1 + Age + ODOS, 
                       data = df,
                       method = "optimal",     
                       ratio = 1)              
summary(match_model)

# matched dataset
matched_data <- match.data(match_model)
# matched_data[matched_data['subclass'] == 1, ] # example

# treated <- matched_data[matched_data$TarsusresectedYesNo == 1, ]
# control <- matched_data[matched_data$TarsusresectedYesNo == 0, ]
treated <- matched_data[matched_data$TarsusresectedYesNo == "Yes", ]
control <- matched_data[matched_data$TarsusresectedYesNo == "No", ]
paired_data <- merge(treated, control, by = "subclass", suffixes = c("_treated", "_control"))


# # Nonparametric Test: Wilcox test
# wilcox.test(MostrecentPostOpMRD1 ~ TarsusresectedYesNo, data = matched_data)
# wilcox.test(treated$MostrecentPostOpMRD1, control$MostrecentPostOpMRD1, paired = TRUE)

# Unpaired
wilcox.test(paired_data$MostrecentPostOpMRD1_treated,paired_data$MostrecentPostOpMRD1_control)

# Paired
wilcox.test(paired_data$MostrecentPostOpMRD1_treated, paired_data$MostrecentPostOpMRD1_control, paired = TRUE)
```
Lower p-value of 0.03532 and 0.02541 with method = "optimal"


```{r}
## Method = "full"

# df$ODOS <- factor(df$ODOS)

# only included vars. with complete data 
match_model <- matchit(TarsusresectedYesNo ~ PreopMRD1 + Age + ODOS + Gender, 
                       data = df,
                       method = "full")              
summary(match_model)

# matched dataset
matched_data <- match.data(match_model)
# matched_data[matched_data['subclass'] == 1, ] # example

# treated <- matched_data[matched_data$TarsusresectedYesNo == 1, ]
# control <- matched_data[matched_data$TarsusresectedYesNo == 0, ]
treated <- matched_data[matched_data$TarsusresectedYesNo == "Yes", ]
control <- matched_data[matched_data$TarsusresectedYesNo == "No", ]
paired_data <- merge(treated, control, by = "subclass", suffixes = c("_treated", "_control"))


# # Nonparametric Test: Wilcox test
# wilcox.test(MostrecentPostOpMRD1 ~ TarsusresectedYesNo, data = matched_data)
# wilcox.test(treated$MostrecentPostOpMRD1, control$MostrecentPostOpMRD1, paired = TRUE)

# Unpaired
wilcox.test(paired_data$MostrecentPostOpMRD1_treated,paired_data$MostrecentPostOpMRD1_control)

# Paired
wilcox.test(paired_data$MostrecentPostOpMRD1_treated, paired_data$MostrecentPostOpMRD1_control, paired = TRUE)
```

Lower p-value of 0.001115 and 0.0005508 with method = "full"


```{r}
## Method = "quick"

# df$ODOS <- factor(df$ODOS)

# only included vars. with complete data 
match_model <- matchit(TarsusresectedYesNo ~ PreopMRD1 + Age + ODOS + Gender, 
                       data = df,
                       method = "quick")              
summary(match_model)

# matched dataset
matched_data <- match.data(match_model)
# matched_data[matched_data['subclass'] == 1, ] # example

# treated <- matched_data[matched_data$TarsusresectedYesNo == 1, ]
# control <- matched_data[matched_data$TarsusresectedYesNo == 0, ]
treated <- matched_data[matched_data$TarsusresectedYesNo == "Yes", ]
control <- matched_data[matched_data$TarsusresectedYesNo == "No", ]
paired_data <- merge(treated, control, by = "subclass", suffixes = c("_treated", "_control"))


# # Nonparametric Test: Wilcox test
# wilcox.test(MostrecentPostOpMRD1 ~ TarsusresectedYesNo, data = matched_data)
# wilcox.test(treated$MostrecentPostOpMRD1, control$MostrecentPostOpMRD1, paired = TRUE)

# Unpaired
wilcox.test(paired_data$MostrecentPostOpMRD1_treated,paired_data$MostrecentPostOpMRD1_control)

# Paired
wilcox.test(paired_data$MostrecentPostOpMRD1_treated, paired_data$MostrecentPostOpMRD1_control, paired = TRUE)
```

Lower p-value of 0.0004758 and 0.0001129 with method = "quick"


```{r}
## Method = "cem"

# df$ODOS <- factor(df$ODOS)

# only included vars. with complete data 
match_model <- matchit(TarsusresectedYesNo ~ PreopMRD1 + Age + ODOS, 
                       data = df,
                       method = "cem")              
summary(match_model)

# matched dataset
matched_data <- match.data(match_model)
# matched_data[matched_data['subclass'] == 1, ] # example

# treated <- matched_data[matched_data$TarsusresectedYesNo == 1, ]
# control <- matched_data[matched_data$TarsusresectedYesNo == 0, ]
treated <- matched_data[matched_data$TarsusresectedYesNo == "Yes", ]
control <- matched_data[matched_data$TarsusresectedYesNo == "No", ]
paired_data <- merge(treated, control, by = "subclass", suffixes = c("_treated", "_control"))


# # Nonparametric Test: Wilcox test
# wilcox.test(MostrecentPostOpMRD1 ~ TarsusresectedYesNo, data = matched_data)
# wilcox.test(treated$MostrecentPostOpMRD1, control$MostrecentPostOpMRD1, paired = TRUE)

# Unpaired
wilcox.test(paired_data$MostrecentPostOpMRD1_treated,paired_data$MostrecentPostOpMRD1_control)

# Paired
wilcox.test(paired_data$MostrecentPostOpMRD1_treated, paired_data$MostrecentPostOpMRD1_control, paired = TRUE)
```

Lower p-value of 0.01727 and 0.003684 with method = "cem"





```{r}
## Method = "genetic"

# only included vars. with complete data 
match_model <- matchit(TarsusresectedYesNo ~ PreopMRD1 + Age + Gender + ODOS + TissueResectedmm, 
                       data = df,
                       method = "genetic",     
                       ratio = 1)              
summary(match_model)

# matched dataset
matched_data <- match.data(match_model)
matched_data[matched_data['subclass'] == 1, ] # example

# Nonparametric Test: Wilcox test
wilcox.test(MostrecentPostOpMRD1 ~ TarsusresectedYesNo, data = matched_data)
wilcox.test(treated$MostrecentPostOpMRD1, control$MostrecentPostOpMRD1, paired = TRUE)


```




## Cj Propensity Matching

```{r}
# Method argument = "optimal"
match_model <- matchit(TarsusresectedYesNo ~ PreopMRD1 + Age + 
                                              ODOS + TissueResectedmm, 
                       data = df,
                       method = "optimal",     
                       ratio = 1)              
summary(match_model)

# matched dataset
matched_data <- match.data(match_model)
matched_data

boxplot(MostrecentPostOpMRD1 ~ TarsusresectedYesNo, data = matched_data)
t.test(MostrecentPostOpMRD1 ~ TarsusresectedYesNo, data = matched_data)

plot(matched_data, type = "density")

```







## Andre Data Cleaning
```{r}
df <- read.csv("new_updated.csv") # from github

# simplify colnames
colnames(df) <- gsub("[[:punct:]]", "", colnames(df))

# change resp var to binary for matching 
df['TarsusresectedYesNo'] <- sapply(df[['TarsusresectedYesNo']], function(x) {ifelse(x == "Y", 1, 0)})

# impute gender by mode
get_mode <- function(x) {
  ux <- na.omit(unique(x))
  ux[which.max(tabulate(match(x, ux)))]
}

mode_value <- get_mode(df[['GenderMale1Female2']])
df$GenderMale1Female2[is.na(df$GenderMale1Female2)] <- mode_value

# impute tissueresected by mean
df$TissueResectedmm[is.na(df$TissueResectedmm)] <- mean(df$TissueResectedmm, na.rm = TRUE)

# for patients with surgery in two eyes, remove the non-tarsectomy
df <- df[!(df$X %in% c(156, 157, 158)), ]

head(df)
```





#1: Andre Propensity Matching (Ratio 1:1)
```{r}
# only included vars. with complete data 
match_model <- matchit(TarsusresectedYesNo ~ PreopMRD1 + Age + GenderMale1Female2 + ODOS + TissueResectedmm, 
                       data = df,
                       method = "nearest",     
                       ratio = 1)              
summary(match_model)

# matched dataset
matched_data <- match.data(match_model)
matched_data[matched_data['subclass'] == 1, ] # example

# balance covariates
plot(match_model, type = "density", interactive = FALSE,
     which.xs = ~ PreopMRD1 + Age + GenderMale1Female2 + ODOS + TissueResectedmm)
summary(match_model)


# boxplot for postopmrd1
boxplot(MostrecentPostOpMRD1 ~ TarsusresectedYesNo, data = matched_data,
        main = "Post-op MRD1 by Tarsectomy with Propensity Score Matching",
        xlab = "Tarsectomy (0 = No, 1 = Yes)",
        ylab = "Post-op MRD1 (mm)",
        col = c("skyblue", "salmon"),
        border = "gray40")

# plot of differences
ggplot(matched_data, aes(x = factor(subclass), y = MostrecentPostOpMRD1, color = factor(TarsusresectedYesNo), group = subclass)) +
  geom_point(size = 2) +
  geom_line() +
  labs(x = "Matched Pair", y = "Post-Op MRD1", color = "Group") +
  scale_color_manual(values = c("blue", "red"), labels = c("Control", "Treatment")) +
  theme_minimal() +
  theme(axis.text.x = element_blank())
```

Parametric Tests
```{r}
# Unpaired T-Test
t.test(MostrecentPostOpMRD1 ~ TarsusresectedYesNo, data = matched_data)

# Paired T-Test
treated <- matched_data[matched_data$TarsusresectedYesNo == 1, ]
control <- matched_data[matched_data$TarsusresectedYesNo == 0, ]
paired_data <- merge(treated, control, by = "subclass", suffixes = c("_treated", "_control"))
t.test(paired_data$MostrecentPostOpMRD1_treated, paired_data$MostrecentPostOpMRD1_control, paired = TRUE)

```

Nonparametric Tests
```{r}
# Unpaired
wilcox.test(treated$MostrecentPostOpMRD1, control$MostrecentPostOpMRD1)

# Paired
wilcox.test(treated$MostrecentPostOpMRD1, control$MostrecentPostOpMRD1, paired = TRUE)
```

#2: Propensity Matching (Ratio 1:2)

```{r}
# only included vars. with complete data 
match_model <- matchit(TarsusresectedYesNo ~ PreopMRD1 + Age + GenderMale1Female2 + ODOS + TissueResectedmm, 
                       data = df,
                       method = "nearest",     
                       ratio = 2)              
summary(match_model)

# matched dataset
matched_data <- match.data(match_model)
matched_data[matched_data['subclass'] == 1, ] # example

# balance covariates
plot(match_model, type = "density", interactive = FALSE,
     which.xs = ~ PreopMRD1 + Age + GenderMale1Female2 + ODOS + TissueResectedmm)
summary(match_model)

# boxplot for postopmrd1
boxplot(MostrecentPostOpMRD1 ~ TarsusresectedYesNo, data = matched_data,
        main = "Post-op MRD1 by Tarsectomy with Propensity Score Matching",
        xlab = "Tarsectomy (0 = No, 1 = Yes)",
        ylab = "Post-op MRD1 (mm)",
        col = c("skyblue", "salmon"),
        border = "gray40")

# plot of differences
ggplot(matched_data, aes(x = factor(subclass), y = MostrecentPostOpMRD1, color = factor(TarsusresectedYesNo), group = subclass)) +
  geom_point(size = 2) +
  geom_line() +
  labs(x = "Matched Pair", y = "Post-Op MRD1", color = "Group") +
  scale_color_manual(values = c("blue", "red"), labels = c("Control", "Treatment")) +
  theme_minimal() +
  theme(axis.text.x = element_blank())
```

Tests
```{r}
# Unpaired T-Test (Parametric)
t.test(MostrecentPostOpMRD1 ~ TarsusresectedYesNo, data = matched_data)

# Unpaired (Nonparametric)
wilcox.test(treated$MostrecentPostOpMRD1, control$MostrecentPostOpMRD1)
```

